A cautionary note on the use of Ornstein Uhlenbeck and other models in macroevolutionary studies

### 2015-04-29 Manuscript/ohyou.tex
Updates and destruction of main text

  Most model-based methods for characterizing trait evolution are based on the Brownian constant variance model \citep[for exceptions see][]{price1997correlated,harvey2000comparative,freckleton2006detecting}. 
  The Brownian model, first applied in a phylogenetic context by \citet{cavalli1967} and to across-species data by \citet{felsenstein1973maximum}, is a simple model of trait evolution in which trait variance accrues as a linear function of time, and makes the prediction that traits of closely-related species are more similar than those of distantly-related ones. 
  The Brownian model has been modified in various ways to account for a suite of ecological and evolutionary processes \citep[e.g.,][]{grafen1989phylogenetic,hansen1997stabilizing,Pagel:1997aa,Pagel:1999aa}. 
  The majority of these involve a transformation of the tree and thereby fitting a model with one or more extra parameters. 
  These modified Brownian models tend to fit better and often have links to process-based interpretations.

  One of the most commonly used Brownian-like models is the Ornstein Uhlenbeck (OU) model. 
  This is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing}. 
  The model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait is drawn towards a fitness optimum on an adaptive landscape. 
  The process operating in comparative data is analogous, although clearly is not stabilizing selection, despite being sometimes referred to as such (see section X).
  % Add section no here
  Unfortunately, although widely used, the properties of the OU model are poorly understood leading to the potential for inappropriate use and misinterpretation of results.

% This paragraph may need modifying once we finish
  In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology.
  We use simulations to demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$ that describes the strength of pull towards a central value (typically referred to as the trait or selective optima). We discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses. 
  We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
  Many of our findings are applicable to other models of evolution (see section X) 
  % Add section no here
  , but we focus on the OU model because of the recent increase in publications using the model and because of the under-appreciated ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter. 
  We are not the first to mention these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.

@@

  % I've left the actual details of the lit review in supp mat because I'm not usre how necessary they are here.
  \section{Uses of OU models}
  The popularity of the OU model has grown extensively in recent years (Fig. \ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material). % Need to add ref to supp mat
  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}. 
  It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilizing selection'', this is inaccurate and misleading. As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche. 
  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optima;][]{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.

@@

  \subsection{Performance of $\alpha$ and Likelihood ratio tests}
    To determine whether $\alpha$ is biased, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}. 
    This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model \citep{}. % Add citations
    We then estimated the rejection rate for each set of simulations, i.e., the proportion of times OU models are favoured relative to a Brownian motion model, even though the data were simulated under Brownian model. 
    This gives us a measure of Type I error rate for the OU model.

@@

  \subsection{Simulating phylogenies and data with error}
    We used the same procedure as above to simulate trait data under a Brownian motion model with known error. 
    % should we say "as above" here or be specific i.e. copy paste from above?
    Specifically, we simulated trees under a Yule model with 25, 50, 100, 150, 200, 500, or 1000 tips and added branch length of (i) 1\%, (ii) 5\% or (iii) 10\% of the tree height to the tips of the simulated trees. 
    We then simulated data under a Brownian model on each tree. 
    All our simulated trees and data are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.

  \subsection{Effects of measurement error on OU model performance}
    We compared the fit of a Brownian model to that of an OU model using likelihood ratio tests as described above, using the original trees without the addition of extra branch length to the tip edges. 
    Our expectation is that the OU model should fit the data better than the Brownian model but the reason for the better fit would be entirely unrelated to any evolutionary process.

@@

\section{Prospects: limitations, recommendations and potential solutions?}
  \subsection{Limitations of the OU model}
  Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways. 
  Transformations of the variance-covariance matrix in the Brownian model are an attractive and computationally simple way to modify the basic model to include evolutionary processes. 
  But, as first pointed out by \citet{grafen1989phylogenetic}, the statistical consequences of these modifications can include biases and problems with interpretation. 
  The results of our simulations illustrate that such biases can occur under conditions that closely match the size and type of datasets that are commonly used (Figure \ref{figure:ntaxa}, Supplementary Table S1).

  In the case of the OU model, there are several limitations worth highlighting:
\begin{enumerate}[(i)]
  \item \textbf{Type I error rates are high when sample size is low.} 
  The results of the simulations indicate that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases.  
  \item \textbf{Likelihood ratio tests are untrustworthy.}
  We would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. 
  Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as MCMC, should be considered. 
  \item \textbf{Measurement error increases Type I error rates.} 
  Our results show that a simple Brownian process can be mistaken for an OU process when a small amount of error is added to the data. 
  The effects of measurement error become more severe with increasing tree size.
\end{enumerate}

  These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution. 
  For example, inference of phylogenetic niche conservatism may be particularly problematic because climatic niche traits can vary greatly over the range of species. 
  Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}. 
  A more parsimonious explanation would be that there is unaccounted intraspecific variation or measurement error in the data.

@@

\subsection{Problems, potential solutions and recommendations}
  Although there are clearly limitations with the OU model, at least some can be dealt with through judicious use of the data, additional analyses and cautious interpretation of model parameters. We discuss the main issues that we have identified and highlight possible solutions below. Broadly, the recommendations can be divided into approaches to identifying the best fitting model and approaches to interpret model parameters.

  \begin{enumerate}[(i)]
    \item \textbf{Tree size} 
    OU models should not be applied to small trees. Our simulations indicate that trees with >200 tips are necessary to obtain acceptable type I error rates. However, we are cautious with recommending a minimum tree size because the performance of the OU model may vary among datasets for reasons other than tree size. We note also that large trees are particularly susceptible to issues arising from unaccounted measurement error in the data.
    \item \textbf{Simulate fitted models and compare these to your empirical results.}
    A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values. These can then be compared to results for your dataset \citep[see][for a related approach]{slater2013robust}. This is important because we have shown that the shape of a phylogeny has consequences for parameter biases and hypothesis tests. Any given tree will therefore generate unique parameter estimates. Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your}. Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias. 
    \item \textbf{Consider plausible alternative hypotheses.}
    Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured. Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
  In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
  Thus this does not reflect any kind of OU process underlying the data. The similarity between Brownian motion and OU models with small $\alpha$ are demonstrated in Figure XX.

  % Rob's old figure 2 back in here with a fairly lengthy figure legend (mainly so we don't need it in the main text)

 To avoid misinterpreting results in these cases, we recommend inspecting the $\alpha$ parameter values as well as looking at which model is favoured. 
  If $\alpha$ is small, then this should not be interpreted as any OU-like process (including phylogenetic niche conservatism etc.). Similarly, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.

  The $\alpha$ parameter ranges from 0 to infinity and recognising small or large values may not be intuitive. $\alpha$ can sometimes more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:

@@

  \noindent
  This is sometimes referred to as a measure of ``phylogenetic inertia'' but we dislike this term because it means different things to different researchers \citep{Losos:2008aa}.
  If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
  We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum''.
  However, if $t_\frac{1}{2}$ is extremely large, it suggests that if an OU process is acting, it is extremely weak, thus should not be interpreted as evidence of any kind of process. 
  As a further note of caution, it is important to recognise that biases in the estimation of $\alpha$ would clearly lead to similarly biases in the phylogenetic half-life.

@@

    \item \textbf{Account for measurement error.} % details
    % This is from the original Discussion

    % I dislike referring to Rob's unpublished work. I think here we need.

    %In terms of the final point, it would be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}. It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}. However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling. This is because the variances will themselves be estimates and subject to error. Treating such variances as error-free can potentially create biases (Freckleton unpub). If the variances are unknown then Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} can be used to estimate the non-Brownian component of the trait variance. However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances. 
    \item \textbf{If possible, include ancillary data.}
    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}
    This is implemented in GEIGER \citep{Harmon:2008aa}.
    A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances (see Appendix \ref{equation:OUcov} depend on both the shared distances between species and the distance of a node to the nearest tip. 
    This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G. Slater unpub.). 
    Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. 
    These implementations should not be used with fossil data.

    \item \textbf{Use a Bayesian approach.} See below. %Or cite our preprint?
\end{enumerate}

\subsection{Bayesian simulations} 
% (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)
    For completeness, we repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. 
    The choice of priors for OU models has not been fully explored. 
    Therefore, we used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20. 
    For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
    We present the results from the exponential prior with mean = 10 in the main text since this is a broad, liberal prior, but we provide results derived from the other priors in the Supplementary Material (Figs. S2-S5). % add links to this
    We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. 
    Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. 
    Multiple independent chains were run for each analysis to ensure convergence was reached. 
    We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}implemented in BayesTraits. 
    The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). 
    Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded) 
    We treated Bayes factors $>$ 2 as evidence favouring the OU model.

\section{Just an OU problem?}
%Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}

@@

\section{Concluding comments/Conclusions}
% Conclusion from original
%In conclusion, the simulations we report highlight that there are some limits to what we can learn from phylogenies and comparative data on extant species. The OU model is a good example of this: large values of $\alpha$ will lead to a loss of history and the dataset will only represent the most recent evolutionary changes. Moreover, different processes can easily yield similar patterns \citep[e.g.,][]{Revell:2008ab} with the consequence that rejection of the Brownian model in favour of another does not necessarily say anything about process. This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}. We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation. The results presented above highlight that in the case of the OU model, this is especially important.

>>>>

  The majority of PCMs use an explicit evolutionary model to characterize trait evolution (Freckleton et al. 2011). Most model-based methods for characterizing trait evolution are based on the Brownian constant variance model \citep[for exceptions see][]{price1997correlated,harvey2000comparative,freckleton2006detecting}. 
  The Brownian model, first applied in a phylogenetic context by \citet{cavalli1967} and to across-species data by \citet{felsenstein1973maximum}, is a simple model of trait evolution in which trait variance accrues as a linear function of time, and makes the prediction that traits of closely-related species are more similar than those of distantly-related ones. 
  The Brownian model has been modified in various ways to account for a suite of ecological and evolutionary processes \citep[e.g.,][]{grafen1989phylogenetic,hansen1997stabilizing,Pagel:1997aa,Pagel:1999aa}. 
  The majority of these involve a transformation of the tree and thereby fitting a model with one or more extra parameters. 
  These modified Brownian models tend to fit better and often have links to process-based interpretations.

  One of the most commonly used Brownian-like models is the Ornstein Uhlenbeck (OU) model. 
  The OU model was introduced to population genetics by \cite{Lande:1976aa} to model stabilizing selection in which the trait is drawn towards a fitness optimum on an adaptive landscape. 
  The process operating in comparative data is analogous to but distinct from stabilizing selection. The phylogenetic OU model is a modification of the Brownian model with an additional parameter $\alpha$ that measures the strength of return towards a theoretical optimum \citep{hansen1997stabilizing} that is shared across a clade or subset of species.
  Although widely used, the properties of the OU model, and other direct extensions of the Brownian model, are poorly understood leading to the potential for inappropriate use and misinterpretation of results.

The popularity of the OU model has grown extensively in recent years (Fig. \ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material). % Need to add ref to supp mat
  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}. 
  It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilizing selection'', this is inaccurate and misleading. As formulated by \citet{hansen1997stabilizing}, a niche has a primary optimum that is the mean of individual species optima for that niche. 
  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optima;][]{hansen2012adaptive} but not as an estimate of stabilizing selection in the population genetics sense.

@@

% This paragraph may need modifying once we finish
  In this paper we present an introduction to the OU model, its general properties and some issues with its use in ecology, evolution and palaeontology.
  We use simulations to demonstrate the inherent bias in estimating the core parameter of the OU model, $\alpha$ that describes the strength of pull towards a central value (typically referred to as the trait or selective optima). We discuss the intricacies of interpreting OU models biologically, and provide advice for appropriate use of OU models in phylogenetic comparative analyses. 
  We also show that very small amounts of intraspecific trait variation (including measurement error) in data can have profound effects on the performance of models. 
  Many of our findings are applicable to other models of evolution (see section X) 
  % Add section no here
  , but we focus on the OU model because of its widespread use and because of the ambiguity in the link between pattern and process when interpreting estimates of the $\alpha$ parameter. 
  We are not the first to describe these problems \citep[e.g.,][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your,hansen2012interpreting,ives2010phylogenetic}, but we hope by summarising the issues and giving recommendations we can help those less familiar with the technical literature.

@@

  \subsection{Performance of $\alpha$ and Likelihood ratio tests}
    To determine whether $\alpha$ is biased, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT \citep{Thomas:2011aa}. 
    This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model \citep{}. % Add citations
    We then estimated the rejection rate of the null (Brownian) model for each set of simulations. 
    This gives us a measure of Type I error rate for the OU model.

@@

    We used the same procedure as above to simulate trait data under a Brownian motion model with known error. 
    % should we say "as above" here or be specific i.e. copy paste from above?
    Specifically, we simulated trees under a Yule model with 25, 50, 100, 150, 200, 500, or 1000 tips and added branch length of (i) 1\%, (ii) 5\% or (iii) 10\% of the tree height to the tips of the simulated trees. 
    We then simulated data under a Brownian model on each tree. 
    All our simulated trees and data are available on GitHub: \href{https://github.com/nhcooper123/OhYou}{github.com/nhcooper123/OhYou}.

    We compared the fit of a Brownian model to that of an OU model using likelihood ratio tests as described above, using the original trees without the addition of extra branch length to the tip edges. 
    Our expectation is that the OU model should fit the data better than the Brownian model because the $\alpha$ parameter should account for much of the error. In this case, rejection of the Brownian model does not represent type I error because it would be quite correct to reject the Brownian. However, the reason for the better fit would be entirely unrelated to any macroevolutionary process.

@@

  \subsection{Limitations of the OU model}
  Although it is possible to create and implement new models for comparative data that encompass a range of processes, we have to beware that such models are statistically complex and may behave in unexpected ways. 
  Transformations of the variance-covariance matrix in the Brownian model are an attractive and computationally simple way to modify the basic model to include evolutionary processes. 
  But, as first pointed out by \citet{grafen1989phylogenetic}, the statistical consequences of these modifications can include biases and problems with interpretation. 
  The results of our simulations illustrate that such biases can occur under conditions that closely match the size and type of datasets that are commonly used (Figure \ref{figure:ntaxa}, Supplementary Table S1).

  In the case of the OU model, there are several limitations worth highlighting:
\begin{enumerate}[(i)]
  \item \textbf{Type I error rates are high when sample size is low.} 
  The results of the simulations indicate that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases.  
  \item \textbf{Likelihood ratio tests are untrustworthy.}
  We would expect these to be approximate because $\alpha$ is bounded and has a non-linear effect on the expected variances. 
  Our simulations indicate that Likelihood ratio tests should not be relied upon for analyses with small sample sizes, and that for robust inference and testing, alternatives, such as MCMC, should be considered. 
  \item \textbf{Measurement error increases Type I error rates.} 
  Our results show that a simple Brownian process can be mistaken for an OU process when a small amount of error is added to the data. 
  The effects of measurement error become more severe with increasing tree size.

  These limitations mean that when evidence for the OU model is found, the results should be interpreted with caution.
\end{enumerate}

  % As dodgy as some of the niche conservatism literature is, I think this paragraph is out of place. I think that is a separate debate and probably not one we want to get into here. 
  % For example, inference of phylogenetic niche conservatism may be particularly problematic because climatic niche traits can vary greatly over the range of species. 
 % Although definitions of phylogenetic niche conservatism vary \citep{Losos:2008aa,Losos:2008ab,Wiens:2008aa,crisp2012phylogenetic} one proposed approach is to fit an OU model and take rejection of BM in favour of OU as evidence of constraints on niche evolution \citep{Wiens:2010aa}. 
 % A more parsimonious explanation would be that there is unaccounted intraspecific variation or measurement error in the data.

@@

\section{Prospects: limitations, recommendations and potential solutions}

  We have highlighted several issues with the OU model. These can be broadly divided into two classes: (i) fitting the model and (ii) interpreting model parameters. These issues can be at least partially addressed with additional or alternative analytical approaches. Below we make recommendations for best practice in fitting model and approaches to interpret model parameters. We highlight areas of future research that may be important in alleviating outstanding issues.

  \subsection{Recommendations for model fitting}
    \begin{enumerate}[(i)]
    \item \textbf{Simulate under the null model} OU models should not be applied to small trees. Our simulations indicate that trees with >200 tips are necessary to obtain acceptable type I error rates. However, we are cautious with recommending a minimum tree size because the performance of the OU model may vary among datasets for reasons other than tree size. We note also that large trees are particularly susceptible to issues arising from unaccounted measurement error in the data. Instead, we suggest using simulations to asses the model fit. Simulating data under the BM model will generate null distributions \citep[e.g.,][]{boettiger2012your} and allow straightforward assessment of model fit, for example by generating appropriate critical values for likelihood ratio tests. %We could potentially add a script/workflow for this

    \item \textbf{Consider Bayesian approaches} An alternative but little explored approach is to use Bayesian methods. Bayesian model fitting has not been widely available for OU model fitting until recently but is now possible in R packages including diversitree and geiger (the Single Stationary Peak model in fitContinuousMCMC) and stand alone software including BayesTraits. %citations needed

	We repeated the Maximum Likelihood simulations in a Bayesian framework implemented in BayesTraits. We determined fit to an OU model using Bayes factors estimated from a stepping stone sampling procedure \citep{xie2010improving}. The marginal likelihoods of the models were calculated using a stepping stone sampler in which fifty stones were drawn from a beta distribution with (alpha = 0.4 and beta = 1). Each stone was sampled for 20000 iterations (with the first 5000 iterations discarded). We treated Bayes factors $>$ 2 as evidence favouring the OU model. We ran the MCMC chains for $1x10^6$ iterations, disregarding the first $1x10^4$ as burn-in. Following burn-in the chains were sampled every 1000 to ensure independence of each consecutive sample. Multiple independent chains were run for each analysis to ensure convergence was reached. An important challenge for Bayesian approaches is selection of appropriate priors. We used three alternative sets of priors on $\alpha$: (i) an exponential distribution with mean = 1; (ii), an exponential distribution with mean = 10, and (iii) a uniform distribution bounded at 0 and 20. For all analyses we used a uniform -100 to 100 prior for $\mu$ and uniform 0 to 100 prior for $\sigma^2$. 
    Table XX shows the results from the exponential prior with mean = 10. This is a broad, liberal prior, but our results are similar regardless of priors. %add links to this
    The Bayesian approach results in highly conservative rejection rates regardless for all tree shapes in the absence of measurement error.  While this is encouraging from the perspective of falsely rejecting the Brownian null model, it is also indicative of potentially low statistical power. The Bayesian approach also appears to more readily handle low levels of measurement error. With 1\% measurement error the Bayesian approach retains acceptable rejection rates (<0.05) for trees of up to 150 tips. However, more error or larger trees result in the frequent rejection of the Brownian model. As noted above, this is not an issue of type I and it is entirely correct that the Brownian model is rejected. In such cases emphasis should shift to how the $\alpha$ parameter is interpreted. At this stage we are cautious in recommending Bayesian methods for fitting OU models but instead suggest that this may be a fruitful area for further investigation. 
    \end{enumerate}

  \subsection{Recommendations for interpreting $\alpha$}    
    \begin{enumerate}[(i)]
    \item \textbf{Consider plausible alternative hypotheses} The effects of measurement in particular suggest that $\alpha$ must always be interpreted with caution. Because the OU model was proposed with evolutionary process in mind these alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ can be overlooked. Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured. Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
  In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
  Thus this does not reflect any kind of OU process underlying the data. The similarity between Brownian motion and OU models with small $\alpha$ are demonstrated in Figure XX. A the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.

    A good strategy for data exploration would be to simulate data under BM and fitted OU models to generate distributions of parameters under known values. These can then be compared to results for your dataset \citep[see][for a related approach]{slater2013robust}. This is important because we have shown that the shape of a phylogeny has consequences for parameter biases and hypothesis tests. Any given tree will therefore generate unique parameter estimates.  Generating data under the fitted OU model will allow an assessment of whether it is possible to retrieve known values, or whether there is evidence of bias.

	\item \textbf{Calculate the phylogenetic half-life} 
  The $\alpha$ parameter ranges from 0 to infinity and recognising small or large values may not be intuitive. $\alpha$ can sometimes more easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:

@@

  \noindent
  If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, evolution towards the optimum trait value is fast, residual phylogenetic correlations are weak, and there is little influence of the past on trait values \citep{hansen1997stabilizing}.
  We would not advise interpreting $t_\frac{1}{2}$ as literally being ``the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum''.
  However, if $t_\frac{1}{2}$ is extremely large, it suggests that if an OU process is acting, it is extremely weak, thus should not be interpreted as evidence of any kind of process. 
  As a further note of caution, it is important to recognise that biases in the estimation of $\alpha$ would clearly lead to similarly biases in the phylogenetic half-life.

@@

    \item \textbf{If possible, include ancillary data.}
    If data on fossils are available then these could be incorporated into the analysis \citep{Slater:2012ab}
    This is implemented in GEIGER \citep{Harmon:2008aa}.
    A caution here is that the OU model for non-ultrametric trees has to be carefully parameterized because for non-ultrametric trees the co-variances (see Appendix \ref{equation:OUcov} depend on both the shared distances between species and the distance of a node to the nearest tip. 
    This creates potential problems in parameterization and in interpretation because the variance-covariance matrix is no longer tree-like (G. Slater unpub.). 
    % is this now in Graham's MEE correction?
    Some current implementations of the OU model are based on transforming the tree directly, rather than transforming the variance co-variance matrix \citep[e.g., MOTMOT;][]{Thomas:2011aa}. 
    These implementations should not be used with fossil data.
   \end{enumerate}

\section{Conclusions and outstanding issues}

A recurring theme from our simulations is that interpretation of $\alpha$ is not straightforward in the presence of measurement error. It would clearly be very useful to have estimates of measurement errors for the species’ observations, however the inclusion of species-specific variances has to be done carefully \citep[e.g.,][]{grafen1989phylogenetic}. Several approaches for accounting for error have been proposed %check references. 
It would be tempting to create a vector of variances that could be combined with V to create an overall combined variance matrix \citep{o2012evolutionary,Harmon:2010aa}. However, if the variances are large, vary among species or are based on small samples, this can create additional complications for the modeling. This is because the variances will themselves be estimates and subject to error. Treating such variances as error-free can potentially create biases (Freckleton unpub). An additional challenge common to (especially large) comparative data sets is that measurement error or intraspecific variation is either not available or is inconsistently reported. One approach if the variances are unknown would be to simultaneously fit Pagel’s $\lambda$ \citep{Pagel:1997aa,Pagel:1999aa} to estimate the non-Brownian component of the trait variance. However estimating this parameter simultaneously with estimating $\alpha$ would have to be undertaken carefully as the two transformations have quite similar effects on the predicted variances.

These issues are not limited to OU models. Any model of trait evolution that attempts to account for non-Brownian components of trait variation is susceptible to being mislead by  measurement error. The fundamental problem is that rejection of the Brownian model in favour of another does not necessarily say anything about process. This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}. We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation.

%Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}

@@

% Conclusion from original
%In conclusion, the simulations we report highlight that there are some limits to what we can learn from phylogenies and comparative data on extant species. The OU model is a good example of this: large values of $\alpha$ will lead to a loss of history and the dataset will only represent the most recent evolutionary changes. Moreover, different processes can easily yield similar patterns \citep[e.g.,][]{Revell:2008ab} with the consequence that rejection of the Brownian model in favour of another does not necessarily say anything about process. This problem can be alleviated to some extent if model comparisons are set in a firm hypothesis testing framework in which alternative hypotheses make clear predictions of emerging patterns that can be unambiguously associated with particular models \citep[e.g.,][]{Cooper:2011aa}. We should not use any statistical model without thinking carefully about the limits in terms of both data and interpretation. The results presented above highlight that in the case of the OU model, this is especially important.
